---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
library(knitr)
opts_chunk$set(echo= TRUE, results="hold")
library(data.table)
library(ggplot2)
library(readr)
library(dplyr)
activity <- read_csv("activity.csv")
colnames(activity) <- c("steps", "date", "interval")
activity$date <- as.Date(activity$date, format= "%Y-%m-%d")
activity$interval<- as.factor(activity$interval)
```

##Check the structure of the data
```{r}
str(activity)
```

## What is mean total number of steps taken per day?
```{r}
steps_per_day <- aggregate(steps ~ date, activity, sum)
head(steps_per_day)
```

###Make a histogram of the total number of steps taken each day
```{r}
ggplot(steps_per_day, aes(steps))+geom_histogram()+labs(title="Steps Taken/Day", x="Number of Steps", y="Times/Day")+theme_classic()
```

###The mean and median total number of steps taken per day
```{r}
steps_mean <- mean(steps_per_day$steps, na.rm = TRUE)
steps_median <- median(steps_per_day$steps, na.rm= TRUE)
```

## What is the average daily activity pattern?

###Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
steps_per_interval <- aggregate(activity$steps, by = list(interval = activity$interval), FUN=mean, na.rm=TRUE)
colnames(steps_per_interval) <- c("interval", "steps")
ggplot(steps_per_interval, aes(x=interval, y=steps)) +  geom_point(color="blue") + labs(title="Average Daily Activity Pattern", x="Interval", y="Number of steps") + theme_classic()
```

###Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
max_interval <- steps_per_interval[which.max(steps_per_interval$steps),]
```

## Imputing missing values

###Calculate and report the total number of missing values in the dataset(i.e. the total number of rows with NAs)
```{r}
missing_vals <- sum(is.na(activity$steps))
```

###Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
```{r}
na_fill <- function(data, pervalue) {
        na_index <- which(is.na(data$steps))
        na_replace <- unlist(lapply(na_index, FUN=function(idx){
                interval = data[idx,]$interval
                pervalue[pervalue$interval == interval,]$steps
        }))
        fill_steps <- data$steps
        fill_steps[na_index] <- na_replace
        fill_steps
}
```

###Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
activity_fill <- data.frame(  
        steps = na_fill(activity, steps_per_interval),  
        date = activity$date,  
        interval = activity$interval)
str(activity_fill)
```

###Make a histogram of the total number of steps taken each day. 
```{r}
fill_steps_per_day <- aggregate(steps ~ date, activity_fill, sum)
colnames(fill_steps_per_day) <- c("date","steps")
ggplot(fill_steps_per_day, aes(x = steps)) + 
geom_histogram() + labs(title="Histogram of Steps Taken per Day", x = "Number of Steps per Day", y = "Number of times in a day(Count)") + theme_classic() 
```

###Calculate and report the mean and median total number of steps taken per day.
```{r}
steps_mean_fill <- mean(fill_steps_per_day$steps, na.rm=TRUE)
steps_median_fill <- median(fill_steps_per_day$steps, na.rm=TRUE)
```

## Are there differences in activity patterns between weekdays and weekends?

###Create a new factor variable in the dataset with two levels – “weekday”and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
weekdays_steps <- function(data) {
    weekdays_steps <- aggregate(data$steps, by=list(interval = data$interval),
                          FUN=mean, na.rm=T)
    # convert to integers for plotting
    weekdays_steps$interval <- 
            as.integer(levels(weekdays_steps$interval)[weekdays_steps$interval])
    colnames(weekdays_steps) <- c("interval", "steps")
    weekdays_steps
}

data_by_weekdays <- function(data) {
    data$weekday <- 
            as.factor(weekdays(data$date)) # weekdays
    weekend_data <- subset(data, weekday %in% c("Saturday","Sunday"))
    weekday_data <- subset(data, !weekday %in% c("Saturday","Sunday"))

    weekend_steps <- weekdays_steps(weekend_data)
    weekday_steps <- weekdays_steps(weekday_data)

    weekend_steps$dayofweek <- rep("weekend", nrow(weekend_steps))
    weekday_steps$dayofweek <- rep("weekday", nrow(weekday_steps))

    data_by_weekdays <- rbind(weekend_steps, weekday_steps)
    data_by_weekdays$dayofweek <- as.factor(data_by_weekdays$dayofweek)
    data_by_weekdays
}

data_weekdays <- data_by_weekdays(activity_fill)
```

###Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r}
ggplot(data_weekdays, aes(x=interval, y=steps)) + geom_line() + facet_wrap(~ dayofweek, nrow=2, ncol=1) + labs(x="Interval", y="Number of steps") +theme_classic()
```






